DSSynth: An Automated Digital Controller Synthesis Tool for Physical Plants

### 2017-06-29 ase2017/paper.tex
ase
Signed-off-by: aabate <aabate@cs.ox.ac.uk>

There are various mathematical representations for control 
systems,  e.g.,  transfer functions and state-space equations (difference or differential equations). 
Here, 
systems with a single input and a single output (SISO) are considered and 
represented by transfer functions or state-space LTI models. LTI models 
are used in several engineering applications to support non-trivial tasks when 
applied to embedded systems (e.g., robot manipulator, power plants 
and guidance system). The following expression presents the general form 
of a discrete plant or a digital controller based on the transfer function representation:
%
\begin{equation}
\small
\label{eq:transferfunction}
H(z)=\frac{B(z)}{A(z)}=\frac{b_{0}+b_{1}z^{-1}+...+b_{M}z^{-M}}{1+a_{1}z^{-1}+...+a_{N}z^{-N}},
\end{equation}
%
\noindent where $z^{-1}$ is called backward-shift operator; $A(z)$ and $B(z)$ are 
the denominator and numerator polynomials, respectively; and $N$ and $M$
represent the denominator and numerator polynomials order, respectively.
%
In state-space format, the digital control system represents the behavior 
of a system via a state evolution equation $x(n+1)$ and an instantaneous 
output equation $y(n)$, as follows:
%
\begin{equation}
\begin{split}
x(n+1) &= A x(n) + B u(n)
\\
y(n) &= C x(n) + D u(n), 
\end{split}\label{eq:ss-example}
\end{equation}

\noindent where $A$, $B$, $C$ and $D$ are matrices that fully specify a digital system. 
%
Controllers synthesis for LTI systems is richly studied 
in the literature~\cite{mazo2010pessoa,DBLP:conf/emsoft/RavanbakhshS16,economakos2016automated}; 
however, there are many challenges when 
digital control system theory is used, mainly due to the finite-word 
length (FWL) effects~\cite{Guang2013, Istepanian2001}, time discretization 
and quantization noise, which are typically introduced by Analogue-to-Digital 
(ADC) and Digital-to-Analogue (DAC) conversion, leading to truncation 
or round-off errors~\cite{astrom1997computer}.

Recently, a new methodology to synthesize digital control systems was proposed, 
named as DSSynth (Digital-System Synthesizer)~\cite{abate2017, abatecav2017}, 
which is based on the Counter-Example Guided Inductive Synthesis 
(CEGIS) approach~\cite{DBLP:conf/asplos/Solar-LezamaTBSS06}. Here, CEGIS is used
as a program synthesis engine that is able to produce digital controllers for highly non-trivial 
control system specifications with a very high degree of automation. Program synthesis engines 
use a specification as the starting point, and subsequently produce a sequence of candidate 
programs from a given template. The candidate programs are iteratively refined to eventually 
satisfy the specification. Modern synthesis engines as the one used here combine automated testing, genetic algorithms 
and SMT-based automated reasoning~\cite{DBLP:journals/corr/AlurFSS16a, DBLP:conf/lpar/DavidKL15}. 
DSSynth synthesizes stable digital controllers using that CEGIS-based approach,
for a given physical plant represented by either transfer functions or state-space equations~\cite{abate2017,abatecav2017}.

Notably, there are toolboxes in MATLAB with functions and scripts to facilitate the 
digital system design and implementation~\cite{matlab-toolbox}. Indeed, there is a MATLAB 
Toolbox named as DSVerifier Toolbox~\cite{issta2017}, which automatically detects 
specific errors related to digital system design ({\it e.g.}, stability, limit cycle and overflow) 
using symbolic model checking techniques based on Boolean Satisfiability (SAT) and 
Satisfiability Modulo Theories (SMT) solvers. Pessoa~\cite{mazo2010pessoa} is another 
toolbox applied to the synthesis of correct-by-design digital controllers. 
In particular, Pessoa is based on approximate bisimulation, which allows control engineers to replace differential equations, 
describing a physical plant, by a respective finite-state machine. However, Pessoa does not take advantage of
recent advances in bit-accurate verification of programs to synthesize controllers w.r.t. FWL effects. MATLAB already offers a tool 
named as Control System Tuner~\cite{autotuner} for automatically designing controllers based on optimization solvers or graphical analysis. 
However, the Control System Tuner is unable to consider FWL effects during the controller synthesis process.
As a result, there is no other MATLAB toolbox to perform digital controller synthesis based on 
program synthesis engines and considering the complete range of approximations, including time discretization, 
quantization effects, and finite-precision arithmetic and its rounding errors.

The present paper addresses this problem and describes a toolbox for the DSSynth tool~\cite{abate2017, abatecav2017} 
in MATLAB's environment named as \tool. The main advantage regarding the use of a MATLAB toolbox 
lies on designing physical plants in MATLAB and then promptly synthesizing a stable and safe digital controller,
considering the complete range of approximations. 
Additionally, when using the \tool, an engineer is able to model a physical plant with MATLAB, 
through transfer-function or state-space representations, considering low-level systems parameters 
(implementation features and numerical format), and then obtain a stable and safe synthesized digital-controller.

@@

The design of our synthesizer consists of two phases, an inductive
synthesis phase and a validation phase, which interact via a finite
set of test vectors that is updated incrementally.  Given a stability
and safety specification, the inductive synthesis procedure tries to
find a candidate solution satisfying the specification for the given set of test inputs.
%
If the synthesis phase succeeds in finding a witness, this witness is
a candidate solution to the full synthesis formula.  We pass this
candidate solution to the validation phase, which checks whether it is
a full solution (i.e., satisfies the specification for all possible
inputs).  If this is the case, then the algorithm terminates.
Otherwise, additional information is provided to the inductive
synthesis phase in the form of a new counterexample that is added to
the set of test inputs and the loop iterates again.

We use two different instantiations of the synthesiser
in order to synthesise stable and safe digital
controllers. The two approaches differ in
the manner in which they check the safety specification
for all reachable states.

@@

The first approach presented in Figure~\ref{fig:CEGIS-precision-increment}
uses a multi-staged verification process. 
Essentially, it starts by devising a
digital controller that stabilizes the system while remaining safe for a
pre-selected time horizon and a single initial state; then, it
employs a multi-staged verification process:
(i) The first verification stage checks that the candidate
solution, which we synthesized to be safe for at least one initial
state, is safe for \emph{all} possible initial states, i.e., does not reach
an unsafe state within $k$ steps.
(ii) The second verification stage 
 restores soundness with respect to the plant's precision
by using interval arithmetic \cite{moore1966interval} to validate the 
operations performed by the previous stage. 
(iii) The third verification stage checks that the current
$k$ is large enough to ensure safety for any $k'{>}k$.  Here, we compute the
completeness threshold $\overline{k}$ for the current candidate controller $K$,
i.e., the number of
iterations required to sufficiently unwind the closed-loop state-space
system such that the boundaries are not violated for any larger number of
iterations, and
check that $k{\geq}\overline{k}$.

@@

Instead of unfolding the dynamics, the second approach shown in Figure~\ref{fig:CEGARIS}
employs {\em abstract acceleration}~\cite{cattaruzza2015unbounded} to
evaluate all possible progressions of the system simultaneously. This eliminates the need for
a completenes threshold since the time horizon is always infinite.
Since this model is based on abstraction, we initially perform the synthesis over a basic
model that abstracts away the complexity of the specification and is only refined to a more
complex model when a counterexample is found. Thus, we may only look at the gain of the
system to synthesise a controller and add further restrictions on specific sections of the reach
tube when these prove to invalidate the solution. The use of Counterexample-Guided 
Abstraction Refinement results in much faster synthesis and verification since it is applied to
both phases, allowing for faster results whilst maintaining soundness for an infinite time horizon.

@@

In step $1$, the user has to define the digital controller representation, 
which can be transfer-function or state-space equations. 
In step $2$, the physical plant (represented by Equations (\ref{eq:transferfunction}) 
and (\ref{eq:ss-example})) must be designed, using the preferred method 
from the control literature~\cite{astrom1997computer}. Finally, in step $3$, 
the numerical representation for the digital controller implementation must be 
defined by the user, {\it i.e.}, the FWL format that includes the number of bits for the integer 
and fractional parts as well as the dynamic range inputs. The result of the manual 
procedure performed by the user is a digital system specification for the given 
physical plant in MATLAB.

After that, the synthesis process starts in step $A$, when \tool obtains the digital system 
specification in MATLAB, and then translates it into an ANSI-C program. 
In step $B$, our CEGIS engine is invoked to synthesize the digital controller w.r.t. the digital system
specification. Finally, in step $C$, the synthesized digital controller considering FWL effects is produced. 
The output generated by \tool is the synthesized digital controller represented either in transfer-function 
or state-space format. The synthesis is considered to be \emph{successful} if a digital controller is correctly 
synthesized w.r.t. FWL effects; otherwise the synthesis is considered to be \emph{failed}, if any parameter is 
incorrectly defined in previous steps, or if there is no solution (digital controller) for the respective physical plant
 given the digital controller word-length.

>>>>

There are various mathematical representations for control 
systems,  e.g.,  transfer functions and state-space equations (difference or differential equations). 
Here, 
systems with a single input and a single output (SISO) are considered and 
represented by transfer functions or state-space LTI models. LTI models 
are used in several engineering applications to support non-trivial tasks when 
applied to embedded systems (e.g., robot manipulator, power plants 
and guidance system). The following expression presents the general form 
of a discrete plant or a digital controller based on the transfer function representation:
%
\begin{equation}
\small
\label{eq:transferfunction}
H(z)=\frac{B(z)}{A(z)}=\frac{b_{0}+b_{1}z^{-1}+...+b_{M}z^{-M}}{1+a_{1}z^{-1}+...+a_{N}z^{-N}},
\end{equation}
%
\noindent where $z^{-1}$ is called the backward-shift operator; $A(z)$ and $B(z)$ are 
the denominator and numerator polynomials; and $N$ and $M$
represent the denominator and numerator polynomials order, respectively.
%
In a state-space form, 
the behavior of a system is represented via a state evolution equation $x(n+1)$ and an instantaneous 
output equation $y(n)$, as follows:
%
\begin{equation}
\begin{split}
x(n+1) &= A x(n) + B u(n)
\\
y(n) &= C x(n) + D u(n), 
\end{split}\label{eq:ss-example}
\end{equation}

\noindent where $A$, $B$, $C$ and $D$ are matrices that fully specify the model. 
%
Controllers synthesis for LTI systems is richly studied 
in the literature~\cite{mazo2010pessoa,DBLP:conf/emsoft/RavanbakhshS16,economakos2016automated}; 
however, there are many challenges when 
digital control system theory is used, mainly due to the finite-word 
length (FWL) effects leading to truncation 
or round-off errors~\cite{Guang2013, Istepanian2001}, time discretization 
and quantization noise, which are typically introduced by Analogue-to-Digital 
(ADC) and Digital-to-Analogue (DAC) conversion~\cite{astrom1997computer}.

Recently, a new methodology to synthesize digital control systems has been proposed, 
named DSSynth (Digital-System Synthesizer)~\cite{abate2017, abatecav2017}, 
which is based on the Counter-Example Guided Inductive Synthesis 
(CEGIS) approach~\cite{DBLP:conf/asplos/Solar-LezamaTBSS06}. Here, CEGIS is used
as a program synthesis engine that is able to generate sound digital controllers for highly non-trivial 
control system specifications with a very high degree of automation. Program synthesis engines 
use a specification as the starting point, and subsequently produce a sequence of candidate 
programs from a given template. The candidate programs are iteratively refined to eventually 
satisfy the specification. Modern synthesis engines such as the one used here combine automated testing, genetic algorithms 
and SMT-based automated reasoning~\cite{DBLP:journals/corr/AlurFSS16a, DBLP:conf/lpar/DavidKL15}. 
DSSynth synthesizes stable digital controllers using the CEGIS-based approach,
for a given physical plant represented by either transfer functions or state-space equations~\cite{abate2017,abatecav2017}.

Notably, there are toolboxes in MATLAB with functions and scripts to facilitate the 
digital system design and implementation~\cite{matlab-toolbox}. Indeed, there is a MATLAB 
Toolbox named as DSVerifier Toolbox~\cite{issta2017}, which automatically detects 
specific errors related to digital system design ({\it e.g.}, stability, limit cycle and overflow) 
using symbolic model checking techniques based on Boolean Satisfiability (SAT) and 
Satisfiability Modulo Theories (SMT) solvers. Pessoa~\cite{mazo2010pessoa} is another 
toolbox applied to the synthesis of correct-by-design digital controllers. 
In particular, Pessoa is based on approximate bisimulation, which allows control engineers to replace differential equations, 
describing a physical plant, by a respective finite-state machine. However, Pessoa does not take advantage of
recent advances in bit-accurate verification of programs to synthesize controllers w.r.t. FWL effects. MATLAB already offers a tool 
named as Control System Tuner~\cite{autotuner} \textcolor{red}{[fix ref]} for automatically designing controllers based on optimization solvers or graphical analysis. 
However, the Control System Tuner is unable to consider FWL effects during the controller synthesis process.
As a result, there is no other MATLAB toolbox to perform digital controller synthesis based on 
program synthesis engines and considering the complete range of approximations, including time discretization, 
quantization effects, and finite-precision arithmetic and its rounding errors.

The present paper addresses this limitation and describes a toolbox for the DSSynth tool~\cite{abate2017, abatecav2017} 
in a MATLAB environment, named \tool. The main advantage for the use of a MATLAB toolbox 
lies on designing physical plants in MATLAB and then promptly synthesizing a stable and safe digital controller,
considering the complete range of approximations. 
Additionally, when using the \tool, an engineer is able to model a physical plant with MATLAB, 
through transfer-function or state-space representations, considering low-level systems parameters 
(implementation features and numerical format), and then to synthesize a stable and safe digital controller.

@@

The design of our synthesizer consists of two phases, an inductive
synthesis phase and a validation phase, which interact via a finite
set of test vectors that is updated incrementally.  Given a stability
and safety specification, the inductive synthesis procedure tries to
find a candidate solution satisfying the specification for the given set of test inputs.
%
If the synthesis phase succeeds in finding a witness, this witness is
a candidate solution to the full synthesis formula.  We pass this
candidate solution to the validation phase, which checks whether it is
a full solution (i.e., satisfies the specification for all possible
inputs).  If this is the case, then the algorithm terminates.
Otherwise, additional information is provided to the inductive
synthesis phase in the form of a new counterexample that is added to
the set of test inputs, and the loop iterates again.

We use two different instantiations of the synthesiser
in order to find stable and safe digital
controllers. The two approaches differ in
the manner in which they check the safety specification
for all reachable states.

@@

The first approach, presented in Figure~\ref{fig:CEGIS-precision-increment}, 
uses a multi-staged verification process. 
It starts by devising a digital controller that stabilizes the model while remaining safe for a
pre-selected time horizon ($k$) and a single initial state; then, it
employs a multi-staged verification process as follows:
(i) The first verification stage checks that the candidate
solution, which we synthesized to be safe for at least one initial
state, is safe for \emph{all} possible initial states, i.e., does not reach
an unsafe state within $k$ steps.
(ii) The second verification stage 
 restores soundness with respect to the plant precision
by using interval arithmetic \cite{moore1966interval} to validate the 
operations performed by the previous stage. 
(iii) The third verification stage checks that the current
$k$ is large enough to ensure safety for any $k'{>}k$.  Here, we compute the
completeness threshold $\overline{k}$ for the current candidate controller $K$,
i.e., the number of
iterations required to sufficiently unwind the closed-loop state-space
model such that the boundaries are not violated for any larger number of
iterations, and
check that $k{\geq}\overline{k}$.

@@

Instead of unfolding the dynamics, the second approach (shown in Figure~\ref{fig:CEGARIS}) 
employs {\em abstract acceleration}~\cite{cattaruzza2015unbounded} to 
simultaneously evaluate all possible progressions of the model. This eliminates the need for
a completeness threshold, since the time horizon considered in abstract acceleration is infinite.
Since this model is based on abstraction, we initially perform the synthesis over a basic
model that abstracts away the complexity of the specification and is only refined to a more
complex model when a counterexample is found. Thus, we may only look at the gain of the
model to synthesise a controller and to add further restrictions on specific sections of the reach
tube when these prove to invalidate the solution. The use of this Counterexample-Guided 
Abstraction-Refinement (CEGAR) results in much faster synthesis and verification since it is applied to
both phases, allowing for faster results whilst maintaining soundness for an infinite time horizon.

@@

In step $1$, the user has to define the digital controller representation, 
which can be a transfer function or a state-space model. 
In step $2$, the physical plant (represented by Equations (\ref{eq:transferfunction}) 
and (\ref{eq:ss-example})) must be designed, using a method selected 
from the control literature~\cite{astrom1997computer}. Finally, in step $3$, 
the numerical representation for the digital controller implementation must be 
defined by the user, {\it i.e.}, the FWL format that includes the number of bits for the integer 
and fractional parts as well as the dynamic range inputs. The result of the manual 
procedure performed by the user is a digital system specification for the given 
physical plant in MATLAB.

After that, the synthesis process starts in step $A$, when \tool obtains the digital system 
specification in MATLAB, and then translates it into an ANSI-C program. 
In step $B$, our CEGIS engine is invoked to synthesize the digital controller w.r.t. the digital system
specification. Finally, in step $C$, the synthesized digital controller considering FWL effects is produced. 
The output generated by \tool is the synthesized digital controller represented either in transfer function form 
or state-space format. The synthesis is considered to be \emph{successful} if a digital controller is correctly 
synthesized w.r.t. FWL effects; otherwise the synthesis is considered to be \emph{failed}, if any parameter is 
incorrectly defined in previous steps, or if there is no solution (digital controller) for the respective physical plant 
given the digital controller word length.
